---
author: Omkar Kulkarni
title: "Assignment 7 Risk Management"
output: html_document
---


```{r include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```







```{r include=FALSE}
# install.packages("knitr")
# install.packages("haven")
# install.packages("dplyr")
# install.packages("lubridate")
# install.packages("PerformanceAnalytics")
# install.packages("ggplot2")
# install.packages("fGarch")
# install.packages("sqldf")

library(haven)
library(dplyr)
library(lubridate)
library(PerformanceAnalytics)
library(ggplot2)
library(fGarch)
library(sqldf)
library(knitr)

```


```{r}
dsfa <- read.csv("C:\\Users\\TRANSFORMER\\Desktop\\Favorites\\Downloads\\dsf_a.csv")

dsf <- dsfa[order(dsfa$PERMNO),]

```



```{r}


#Setting seed and finding a random year between 1980 and 2010
set.seed(903467583)
startingYear = floor(runif(1,min=1980,max = 2010))

#Subsetting data for the starting year
dsf_startingYear <- dsf[which(year(dsf$DATE)==startingYear), ]

#Getting list of 100 random firms
set.seed(903467583)
lstPERMNO <- sample(unique(dsf_startingYear$PERMNO),100)

#Subsetting six years of data based on the starting year obtained from sampling
masterdata_6_years <- dsf[ which(year(dsf$DATE)>=startingYear & year(dsf$DATE) <= (startingYear+6) & dsf$PERMNO %in% lstPERMNO), ]

#Subsetting data from 2000 to 2010
masterdata_10_years <- dsf[ which(year(dsf$DATE)>=2000 & year(dsf$DATE) <= 2010 & dsf$PERMNO %in% lstPERMNO), ]

#Finding 1 random PERMNO
set.seed(903467583)
tempPERMNO <- sample(unique(dsf_startingYear$PERMNO),1)

#Subsetting data for the above sampled PERMNO
temp_masterdata_6_years <- dsf[ which(year(dsf$DATE)>=startingYear & year(dsf$DATE) <= (startingYear+6) & dsf$PERMNO == tempPERMNO), ]

#Calculating VaR and ES by functions
VAR_by_function = VaR(temp_masterdata_6_years$RET,p = 0.95, method = "historical")
ES_by_function = ES(temp_masterdata_6_years$RET,p = 0.95, method = "historical")

#Calculating VaR and ES through quantile and mean functions
VAR_by_quantile = quantile(temp_masterdata_6_years$RET,probs = 0.05, na.rm = TRUE)
ES_by_mean = mean(temp_masterdata_6_years[which(temp_masterdata_6_years$RET <= VAR_by_quantile),]$RET, na.rm = TRUE)

#Initializing the initial Investment
initialInvestment <- 1000000

```

```{r}

#Task 1

#Calculating the VaR and ES for 100 PERMNOs in the six years
task1Data <- masterdata_6_years %>%
  group_by(PERMNO) %>%
  summarize(VaR = VaR(RET,p = 0.95, method = "historical"),
            VaR_Dollar_Value = VaR * initialInvestment,
            ES = ES(RET,p = 0.95, method = "historical"),
            ES_DOllar_Value = ES * initialInvestment)

#Calculating VaR and ES for portfolio as a whole in the six years

portfolioData_6_years <- masterdata_6_years %>%
  group_by(DATE) %>%
  summarize(Return = mean(RET,na.rm = TRUE))

Portfolio_VaR_6_years = VaR(portfolioData_6_years$Return,p=0.95,method = "historical")
Portfolio_VaR_Dollar_Value_6_years = Portfolio_VaR_6_years * 100 * initialInvestment
Portfolio_ES_6_years = ES(portfolioData_6_years$Return,p=0.95,method = "historical")
Portfolio_ES_Dollar_Value_6_years = Portfolio_ES_6_years * 100 * initialInvestment

#Task 2

#Calculating the VaR and ES for PERMNOs between 2000 and 2010
task2Data <- masterdata_10_years %>%
  group_by(PERMNO) %>%
  summarize(VaR = VaR(RET,p = 0.95, method = "historical"),
            VaR_Dollar_Value = VaR * initialInvestment,
            ES = ES(RET,p = 0.95, method = "historical"),
            ES_DOllar_Value = ES * initialInvestment)

#Calculating VaR and ES for portfolio as a whole between 2000 and 2010

portfolioData_10_years <- masterdata_10_years %>%
  group_by(DATE) %>%
  summarize(Return = mean(RET,na.rm = TRUE))

Portfolio_VaR_10_years = VaR(portfolioData_10_years$Return,p=0.95,method = "historical")
Portfolio_VaR_Dollar_Value_10_years = Portfolio_VaR_10_years * 46 * initialInvestment
Portfolio_ES_10_years = ES(portfolioData_10_years$Return,p=0.95,method = "historical")
Portfolio_ES_Dollar_Value_10_years = Portfolio_ES_10_years * 46 * initialInvestment

#Task 3

#Initializing lambda
lambda <- 0.94

#Calculating Initial Variances for the firms
initial_variance_data_annualized <- masterdata_6_years %>%
  group_by(PERMNO) %>%
  summarize(Variance_Annualized = var(RET, na.rm = TRUE) * 252)

#Initializing blank data frame
variances_data <- data.frame(PERMNO=NULL, DATE = NULL, RET = NULL, Variance = NULL, stringsAsFactors = FALSE)

#Calculating variances for all firms using the RiskMetrics model to calculate variance
for(tempPERMNO in lstPERMNO){
  
  tempData <- masterdata_10_years[ which(masterdata_10_years$PERMNO==tempPERMNO), ]
  tempData <- tempData[complete.cases(tempData),]
  if(nrow(tempData) > 0)
  {
    tempData$Variance <- NA
    tempData$Variance[1] = initial_variance_data_annualized[ which(initial_variance_data_annualized$PERMNO==tempPERMNO), ]$Variance_Annualized
    
    for(i in 2:nrow(tempData)){
      tempData$Variance[i] = (lambda * tempData$Variance[i-1]/252 + (1 - lambda) * tempData$RET[i-1]**2) *252
    }
    
    temp_variance_data <- data.frame(PERMNO = c(tempPERMNO) , DATE = tempData$DATE, RET = tempData$RET, Variance = tempData$Variance, stringsAsFactors = FALSE)
    
    variances_data <- rbind(variances_data,temp_variance_data)
    
    }
  
  
}


#Task 4

#Initializing blank data frame
garch_coeff_data <- data.frame(PERMNO = NULL, Alpha = NULL, Beta = NULL, Omega = NULL, stringsAsFactors = FALSE)

#Calculating alpha, beta and omega for all 100 PERMNOs
for(tempPERMNO in lstPERMNO){
  
  tempData <- masterdata_6_years[ which(masterdata_6_years$PERMNO==tempPERMNO), ]
  tempData <- tempData[complete.cases(tempData),]
  if(nrow(tempData) > 0)
  {
    garch_model <- garchFit(formula = ~ garch(1, 1), data = tempData$RET)
    alpha <- coef(garch_model)[3]
    beta <- coef(garch_model)[4]
    omega <- coef(garch_model)[2]
    
    temp_coeff_data <- data.frame(PERMNO = c(tempPERMNO), Alpha = c(alpha), Beta = c(beta), Omega = c(omega), stringsAsFactors = FALSE)
    
    garch_coeff_data <- rbind(garch_coeff_data,temp_coeff_data)
  }
  
}

rownames(garch_coeff_data) <- NULL


#Task 5

#Initializing blank data frame
garch_variances_data <- data.frame(PERMNO=NULL, DATE = NULL, RET = NULL, garchVariance = NULL, stringsAsFactors = FALSE)

#Calculating variances for all firms using the GARCH model to calculate variance
for(tempPERMNO in lstPERMNO){
  
  tempData <- masterdata_10_years[ which(masterdata_10_years$PERMNO==tempPERMNO), ]
  tempData <- tempData[complete.cases(tempData),]
  if(nrow(tempData) >0)
  {
    
    tempData$garchVariance <- NA
    tempData$garchVariance[1] = initial_variance_data_annualized[ which(initial_variance_data_annualized$PERMNO==tempPERMNO), ]$Variance_Annualized
    
    tempAlpha <- garch_coeff_data[ which(garch_coeff_data$PERMNO==tempPERMNO), ]$Alpha
    tempBeta <- garch_coeff_data[ which(garch_coeff_data$PERMNO==tempPERMNO), ]$Beta
    tempOmega <- garch_coeff_data[ which(garch_coeff_data$PERMNO==tempPERMNO), ]$Omega
    
    for(i in 2:nrow(tempData)){
      tempData$garchVariance[i] = (tempOmega + tempBeta * tempData$garchVariance[i-1]/252 + tempAlpha * tempData$RET[i-1]**2) * 252
    }
    
    temp_garch_variance_data <- data.frame(PERMNO = c(tempPERMNO) , DATE = tempData$DATE, RET = tempData$RET, garchVariance = tempData$garchVariance, stringsAsFactors = FALSE)
    
    garch_variances_data <- rbind(garch_variances_data,temp_garch_variance_data)
    
    
  }
  
  
}

```

Random Starting Year Generated: 1990  

```{r echo=FALSE, warning=FALSE}

print(paste("VaR generated using the VaR function:",VAR_by_function))
print(paste("VaR generated using the quantile function:",VAR_by_quantile))
print(paste("ES generated using the ES function:",ES_by_function))
print(paste("ES generated using the mean function:",ES_by_mean))

```



#Task 1 - VaR and Expected Shortfall for stocks between 1990 to 1996

##Return Distribution

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}

qplot(masterdata_6_years$RET, geom="histogram", binwidth = 0.005, main = "Return Distribution : 1990-1996", 
      xlab = "Return", ylab = "Frequency", fill=I("green"), col=I("green"), alpha=I(.2)) + xlim(-0.25, 0.25) +
  stat_function(fun = function(x, mean, sd, n, bw){dnorm(x = x, mean = mean, sd = sd) * n * bw}, 
                args = c(mean = mean(masterdata_6_years$RET), sd = sd(masterdata_6_years$RET), n = nrow(masterdata_6_years), bw = 0.005))

```



##Descriptive Statistics of Returns

```{r echo=FALSE, warning=FALSE}

summary_masterdata_6_years = summary(masterdata_6_years$RET)
print("Summary of Returns")
print(summary_masterdata_6_years)


```

##Portfolio Return Distribution

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}

qplot(portfolioData_6_years$Return, geom="histogram", binwidth = 0.005, main = "Return Distribution : 1990-1996", 
      xlab = "Return", ylab = "Frequency", fill=I("green"), col=I("green"), alpha=I(.2)) + xlim(-0.25, 0.25) +
  stat_function(fun = function(x, mean, sd, n, bw){dnorm(x = x, mean = mean, sd = sd) * n * bw}, 
                args = c(mean = mean(portfolioData_6_years$Return), sd = sd(portfolioData_6_years$Return), n = nrow(portfolioData_6_years), bw = 0.005))

```


##VaR and Expected Shortfall

```{r echo=FALSE, warning=FALSE}


print(paste("VaR of the portfolio as a whole:",Portfolio_VaR_6_years))
print(paste("$VaR of the portfolio as a whole:",Portfolio_VaR_Dollar_Value_6_years))
print(paste("ES of the portfolio as a whole:",Portfolio_ES_6_years))
print(paste("$ES of the portfolio as a whole:",Portfolio_ES_Dollar_Value_6_years))

```

##VaR and Expected Shortfall Distribution

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}

qplot(task1Data$VaR, geom="histogram", binwidth = 0.0025, main = "Value At Risk Distribution : 1990-1996", 
      xlab = "Value At Risk", ylab = "Frequency", fill=I("green"), col=I("green"), alpha=I(.2)) + xlim(-0.25, 0.25) +
  stat_function(fun = function(x, mean, sd, n, bw){dnorm(x = x, mean = mean, sd = sd) * n * bw}, 
                args = c(mean = mean(task1Data$VaR), sd = sd(task1Data$VaR), n = nrow(task1Data), bw = 0.0025))

qplot(task1Data$ES, geom="histogram", binwidth = 0.0025, main = "Histogram for Expected Shortfall : 1990-1996", 
      xlab = "Expected Shortfall", ylab = "Frequency", fill=I("green"), col=I("green"), alpha=I(.2)) + xlim(-0.25, 0.25) +
  stat_function(fun = function(x, mean, sd, n, bw){dnorm(x = x, mean = mean, sd = sd) * n * bw}, 
                args = c(mean = mean(task1Data$ES), sd = sd(task1Data$ES), n = nrow(task1Data), bw = 0.0025))

```


##Descriptive Statistics

```{r echo=FALSE, warning=FALSE}

summary_task1Data = summary(task1Data)
print(summary_task1Data)

```


#Task 2 - VaR and Expected Shortfall for stocks between 2000 to 2010

##Return Distribution

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}

qplot(masterdata_10_years$RET, geom="histogram", binwidth = 0.005, main = "Return Distribution : 2000-2010", 
      xlab = "Return", ylab = "Frequency", fill=I("green"), col=I("green"), alpha=I(.2)) + xlim(-0.25, 0.25) +
  stat_function(fun = function(x, mean, sd, n, bw){dnorm(x = x, mean = mean, sd = sd) * n * bw}, 
                args = c(mean = mean(masterdata_10_years$RET), sd = sd(masterdata_10_years$RET), n = nrow(masterdata_10_years), bw = 0.005))

```



##Descriptive Statistics of Returns

```{r echo=FALSE, warning=FALSE}

summary_masterdata_10_years = summary(masterdata_10_years$RET)
print("Summary of Returns")
print(summary_masterdata_10_years)


```

##Portfolio Return Distribution

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}

qplot(portfolioData_10_years$Return, geom="histogram", binwidth = 0.005, main = "Return Distribution : 1990-1996", 
      xlab = "Return", ylab = "Frequency", fill=I("green"), col=I("green"), alpha=I(.2)) + xlim(-0.25, 0.25) +
  stat_function(fun = function(x, mean, sd, n, bw){dnorm(x = x, mean = mean, sd = sd) * n * bw}, 
                args = c(mean = mean(portfolioData_10_years$Return), sd = sd(portfolioData_10_years$Return), n = nrow(portfolioData_10_years), bw = 0.005))

```


##VaR and Expected Shortfall

```{r echo=FALSE, warning=FALSE}


print(paste("VaR of the portfolio as a whole:",Portfolio_VaR_10_years))
print(paste("$VaR of the portfolio as a whole:",Portfolio_VaR_Dollar_Value_10_years))
print(paste("ES of the portfolio as a whole:",Portfolio_ES_10_years))
print(paste("$ES of the portfolio as a whole:",Portfolio_ES_Dollar_Value_10_years))

```

##VaR and Expected Shortfall Distribution

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}

qplot(task2Data$VaR, geom="histogram", binwidth = 0.0025, main = "Value At Risk Distribution : 2000-2010", 
      xlab = "Value At Risk", ylab = "Frequency", fill=I("green"), col=I("green"), alpha=I(.2)) + xlim(-0.25, 0.25) +
  stat_function(fun = function(x, mean, sd, n, bw){dnorm(x = x, mean = mean, sd = sd) * n * bw}, 
                args = c(mean = mean(task2Data$VaR), sd = sd(task2Data$VaR), n = nrow(task2Data), bw = 0.0025))

qplot(task2Data$ES, geom="histogram", binwidth = 0.0025, main = "Histogram for Expected Shortfall : 2000-2010", 
      xlab = "Expected Shortfall", ylab = "Frequency", fill=I("green"), col=I("green"), alpha=I(.2)) + xlim(-0.25, 0.25) +
  stat_function(fun = function(x, mean, sd, n, bw){dnorm(x = x, mean = mean, sd = sd) * n * bw}, 
                args = c(mean = mean(task2Data$ES), sd = sd(task2Data$ES), n = nrow(task2Data), bw = 0.0025))

```


##Descriptive Statistics

```{r echo=FALSE, warning=FALSE}

summary_task2Data = summary(task2Data)
print(summary_task2Data)

```



\newpage

#Task 3 - Simple Risk Metrics Model  


##Variance plot for PERMNO 52090 between 2000 to 2010  

Folowing is the plot generated for PERMNO 52090 whose variances have been calculated using JP Morgan's RiskMetrics variance model: 

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}


  
  temp_variance_data <- variances_data[ which(variances_data$PERMNO==52090), ]
    

  temp_variance_data <- temp_variance_data%>%
    mutate(DATE= as.Date(temp_variance_data$DATE,"%Y-%m-%d"))
  
  plot = ggplot(data = temp_variance_data, aes(x=DATE,y=Variance, group =1)) +
      geom_line(color='steelblue') +
      ylab(label="Variance") +
      xlab(label="DATE")+
    ggtitle(paste("Variance of 52090 over time", sep=" "))+
    scale_x_date(date_breaks = "1 year",date_labels =  "%Y")
  
  print(plot)
  

```


#Task 4 & 5 - GARCH Model 

##Task 4 - Alphas and Betas for 100 PERMNOs

```{r echo=FALSE, warning=FALSE}

print("GARCH Coefficients - Alpha, Beta, and Omega ")

```

\newpage

##Task 5 - GARCH Variance Plot for PERMNO 52090 between 2000 to 2010  

Folowing is the plot generated for the same firm whose variances have been calculated using GARCH(1,1) model: 

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}


  
  temp_garch_variance_data <- garch_variances_data[ which(garch_variances_data$PERMNO==52090), ]

  temp_garch_variance_data <- temp_garch_variance_data%>%
    mutate(DATE= as.Date(temp_garch_variance_data$DATE,"%Y-%m-%d"))
  
  plot = ggplot(data = temp_garch_variance_data, aes(x=DATE,y=garchVariance, group =1)) +
      geom_line(color='steelblue') +
      ylab(label="Variance") +
      xlab(label="DATE")+
    ggtitle(paste("GARCH Variance of 52090 over time", sep=" "))+
    scale_x_date(date_breaks = "1 year",date_labels =  "%Y")
  
  print(plot)


```

##Analysis  

We can see that both models to estimate volatility give us similar graphs for the same firm. In both graphs, we see a huge spike in volatility after 2008 which could be explained by the financial crisis. 


\newpage

#Appendix

##Variance plots for all PERMNOs between 2000 to 2010

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}

lstPERMNO <- unique(variances_data$PERMNO)

for(tempPERMNO in lstPERMNO){
  
  temp_variance_data <- variances_data[ which(variances_data$PERMNO==tempPERMNO), ]
  
  temp_variance_data <- temp_variance_data%>%
    mutate(DATE= as.Date(temp_variance_data$DATE,"%Y-%m-%d"))
  
  
  plot = ggplot(temp_variance_data, aes(x=DATE, color = 'red')) +
      geom_line(aes(y=Variance)) +
      ylab(label="Variance") +
      xlab(label="DATE")+
    ggtitle(paste("Variance of",tempPERMNO,"over time", sep=" "))
  
  print(plot)
  
}

```

##GARCH Variance plots for all PERMNOs between 2000 to 2010

```{r echo=FALSE, fig.height=4, fig.width=8, warning=FALSE}

lstPERMNO <- unique(garch_variances_data$PERMNO)

for(tempPERMNO in lstPERMNO){
  
  temp_garch_variance_data <- garch_variances_data[ which(garch_variances_data$PERMNO==tempPERMNO), ]
  
  temp_garch_variance_data <- temp_garch_variance_data%>%
    mutate(DATE= as.Date(temp_garch_variance_data$DATE,"%Y-%m-%d"))
  
  
  plot = ggplot(temp_garch_variance_data, aes(x=DATE)) +
      geom_line(aes(y=garchVariance), color = 'darkgreen') +
      ylab(label="Variance") +
      xlab(label="DATE")+
    ggtitle(paste("GARCH Variance of",tempPERMNO,"over time", sep=" "))
  
  print(plot)
  
}

```